---
title: "Linear Algebra: wk3 rref"
author: "Bill Chung"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(far)
library(MASS)
library(pracma)
```

# Concepts

## Space, subspace

- Domain, codomain (Range, C($mathbb{A}$))

## Domain, codomain, Range

- You will see the following notation from time to time

$$T: R^n \rightarrow R^m $$

- the above notation is saying that `matrix T will be used to multiply vector with size of n and the resulting vector will have size m`

- And we will get into the details later.

- Vector resize within a space which consist of so many subspaces.

- When you put vectors into a matrix, you get two space, I call them `input` and `output space`.  Input space can be divided into `row space` and `nullspace`, and `output space` can be divided into `column space` and `left null space`

- Think of domain as `row space` and `codomain` as `output space` and `range` as `column space`

## Rank nullity theorem

If $\mathbb{A}$ has $n$ columns, then $\text{Rank}(\mathbb{A}) + \text{dim Nul}(\mathbb{A}) = n$

- see page 156 for the invertible matrix theorem (continued)

## Invertible Linear Transformation

- A linear transformation $\mathbb{T}: R^n \rightarrow R^n$ is said to be `invertible` if there exists a funciton $\mathbb{S}: R^n \rightarrow R^n$ such that 

$$\begin{aligned}
\mathbb{S}(\mathbb{T}(\vec{x}))= \vec{x} \text{ for all }\vec{x}\text{ in } R^n \\
\mathbb{T}(\mathbb{S}(\vec{x}))= \vec{x} \text{ for all }\vec{x}\text{ in } R^n \\
\end{aligned}$$
<br><br>
where dim($\mathbb{A}$) = n by n, $\vec{x}, \vec{b} \in R^n$
- $\mathbb{A}$ is the standard matrix for $\mathbb{T}$
$$\mathbb{A}\vec{x}=\vec{b}$$

- $\mathbb{A}^{-1}$ is the standard matrix for $\mathbb{S}$
$$\mathbb{A}^{-1}\vec{b}=\vec{x}$$

```{r}
r1 <- c(0,1,-4)
r2 <- c(2,-3,2)
r3 <- c(5,-8,9)
A <- rbind(r1,r2,r3)
print(A)
print(rref(A))
```
- Above matrix is invertible matrix based on `rref()`
- Inverse transformation `undo` the transformation

```{r}
x <- c(3,6,9)

#to use the same notation
T <- A
b<- T%*%x
print("Before the transformation")
print(x)
print("After the transformaiton")
print(T%*%x)
```

## When $\mathbb{A}$ is not a square matrix

- With respect to $\mathbb{A}$, $\vec{b}$ is in your `range` and $\vec{x}$ is in `domain`.

- $\mathbb{A}$ transform vectors in domain to range.
- $\mathbb{A}^{-1}$ can transform values in range back to domain, when the $\mathbb{A}$ involved 1-1 transformation.

```{r}
#chapter 1.1 example 3

r1 <- c(0,3,-6,6,4,-5)
r2 <- c(3,-7,8,-5,8,9)
r3 <- c(3,-9,12,-9,6,15)

A <- rbind(r1,r2,r3)
print(A)
Rank(A)
rref(A)

B <- A[,c(1,2,5)]
print(B)
inv(B)
```

## Exercise.  

- Get 5 matrices from class 
- Given a matrix and a vector
  - Show different ways of spanning the given vector

```{r}
print(A)

b <- c(1,5,9)

Rank(A)

H <- cbind(A,b)
Rank(A)

rref(A)

B  <- A[,c(1,2,5)]
print(B)

x_B<- inv(B)%*%b
x_B
B%*%x_B
dim(A)
A%*%c(-38,-9,0,0,7,0)
```


```{r}
r1 <- c(0,3,-6,6,4,-5)
r2 <- c(3,-7,8,-5,8,9)
r3 <- c(3,-9,12,-9,6,15)
r4 <- c(1,3,5,6,7,-1,10)

A <- rbind(r1,r2,r3,r4)
print(A)
dim(A)
Rank(A)

rref(t(A))

rref(A)
B <- A[,c(1,2,3,5)]
Rank(B)
dim(B)
inv(B)
cond(B)

b <- c(0.1, -1.5, 1,-1000)

x_B <- inv(B)%*%b

B%*%x_B

```




  
  
  
  
  
  
