---
title: 'Linear Algebra: wk4 Four Subspaces'
author: "Bill Chung"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(far)
library(MASS)
library(pracma)
```

# Fundamental four subspaces of Matrix

- $R(\mathbb{A}), N(\mathbb{A}), C(\mathbb{A}), N(\mathbb{A}^T)$

## Exercise.  

- Get 5 matrices from class 
- Given a matrix and a vector, 

(1) find out where the vector lives
  - Space and subspace
  
(2) basis of the subspace
(3) Provide a vector that is not in the span of these two subspaces

# Finding a basis of a subspace

## Column space basis

### use rref($\mathbb{A})

```{r}
A = matrix(c(-3,6,-1,1,-7,1,-2,2,3,-1,2,-4,5,8,-4),nrow=3,ncol=5,byrow=TRUE)
print(A)
rref(A)
```

### use orth()

```{r}
C_A = orth(A)
C_A
```

## Left Nullspace($\mathbb{A}^T$)

```{r}
N_AT <- null(t(A))
N_AT
```

## Basis for output space

```{r}
OUT <- cbind(C_A, N_AT)
rref(OUT)
```

## Row space basis 

### Using transformation


### using orth()

```{r}
C_AT <- orth(t(A))
C_AT
```

## Nullspace basis

- Suppose $T(\vec{x}) = A\vec{x}$, then the ```kernel``` or null space of such T can be found as below.

### using nullspace()

```{r}
N_A <- nullspace(A)
N_A
```

## Basis spanning the input space

```{r}
IN <- cbind(C_AT, N_A)
IN
```


- Suppose we have $\vec{H}=[a-3b,b-a,a,b]^T$, this can be written as linear combination of two vectors $a\vec{v}_1$ and $b\vec{v}_2$ where $\vec{v}_1 = [1,-1,1,0]$ and $\vec{v}_2 = [-3,1,0,1]$.

- This is very useful technique of expressing a subspace of $\vec{H}$ as the linear combination of some small collectoin of vectors.

- Subspace of $\vec{H} \in$ Span{$\vec{v}_1, \vec{v}_2$}

## How to find the basis of null space

- Step 1: Given $\mathbb{A}$, find its `rref`
- Step 2: Solve for $\vec{x}$ in $\mathbb{A}\vec{x}=\vec{0}$
- Step 3: express $\vec{x}$ as linear combination of smaller vectors.
- Step 4: identify basis spanning the null space

```{r}
r1 <- c(-3,6,-1,1,-7)
r2 <- c(1,-2,2,3,-1)
r3 <- c(2,-4,5,8,-4)

A <- rbind(r1,r2,r3)

rref(A)
```

```{r}
n1 <- c(2,1,0,0,0)
n2 <- c(1,0,-2,1,0)
n3 <- c(-3,0,2,0,1)

#########################################
# Any vector in the null space with A
#########################################

print(A%*%n1)
print(A%*%(n1+n2+n3))
print(round(A%*%(100*n1+0.1*n2-305*n3),3))
```

# Concept check questions

- What is the relationship between $C(A)$ and $N(A^T)$?
- What is the relationship between $R(A)$ and $N(A)$?
- What is the relationship between $C(A)$ and $R(A)$?
- What is the relationship between $N(A)$ and $N(A^T)$?
- If the basis spanning $C(A)$ are given, can you find out the basis
spanning $N(A^T)$?
- If the basis spanning $R(A)$ are given, can you find out the basis
spanning $N(A^T)$?

