---
title: "wk7"
author: "Bill Chung"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```


```{r include=FALSE}
library(far)
library(MASS)
library(pracma)
```

## I want to test this

- Test once more 

# Solving systems of equations

- by comparing the rank of $\mathbb{A}$ and the augmented matrix, we can find out if the systems of equations have
  - exactly one solutions, (this is case when $\mathbb{A}$ is ```invertible```)
  - many solutions, (the solution set exist in ```hyperplane``` which is ```affine nullspace of $A$```)
  - no solutions (rank of the ```argumented matrix``` is greater than rank of $A$)
  
 
## General approach


$$\begin{aligned}
\mathbb{A}\vec{x} &= \vec{b}\\
\mathbb{B}\vec{x}_B+\mathbb{N}\vec{x}_N &= \vec{b} \\
\mathbb{B}^T\mathbb{B}\vec{x}_B &= \mathbb{B}^T\vec{b} - \mathbb{B}^T\mathbb{N}\vec{x}_N \\
\vec{x}_B &= \big(\mathbb{B}^T\mathbb{B}\big)^{-1}\mathbb{B}^T\vec{b} - \big(\mathbb{B}^T\mathbb{B}\big)^{-1}\mathbb{B}^T\mathbb{N}\vec{x}_N
\end{aligned}$$


where <br> 
$\mathbb{A} = [\mathbb{B}|\mathbb{N}]$ <br>
$\mathbb{B}$ is matrix containing independent column vectors <br>
$\mathbb{N}$ is matrix containing dependent column vectors

## When RHS is not in the span of the column space

$$\hat{\vec{b}} \neq \vec{b}$$ 

$\hat{\vec{b}} \in C(\mathbb{A})$, where $C(\mathbb{A})$ is a column space of $\mathbb{A}$, among all the vectors in $C(\mathbb{A})$.

$$\vec{b} = B_{C(A)}\vec{x}_{C(A)} + B_{N(A^T)}\vec{x}_{N(A^T)}$$

where <br> 
$B_{C(A)}$ and $B_{N(A^T)}$ are the basis of $C(A)$ and $N(A^T)$ <br>
$\vec{x}_{C(A)}$ and $\vec{x}_{N(A^T)}$ are the coordinate of the corresponding basis<br> <br>
Subsituting $B_{C(A)}$ as $B_{C(A)}$ and $\vec{x}_{C(A)}$ as $\vec{x}_B$, the above equation can be written as the following: 

$$B^T\vec{b} = B^TB\vec{x}_{B}$$

$$\vec{x}_B = (B^TB)^{-1}B^T\vec{b}$$

<br> 

and $(B^TB)^{-1}B^T$ is called ```projection matrix of C(A)``` where $B$ are the basis of $C(A)$

$$\vec{b} = B_{C(A)}\vec{x}_{C(A)} + B_{N(A^T)}\vec{x}_{N(A^T)}$$

where <br> 
$B_{C(A)}$ and $B_{N(A^T)}$ are the basis of $C(A)$ and $N(A^T)$ <br>
$\vec{x}_{C(A)}$ and $\vec{x}_{N(A^T)}$ are the coordinate of the corresponding basis<br> <br>
Subsituting $B_{C(A)}$ as $B_{C(A)}$ and $\vec{x}_{C(A)}$ as $\vec{x}_B$, the above equation can be written as the following: 

$$B^T\vec{b} = B^TB\vec{x}_{B}$$

$$\vec{x}_B = (B^TB)^{-1}B^T\vec{b}$$

<br> 

and $(B^TB)^{-1}B^T$ is called ```projection matrix of C(A)``` where $B$ are the basis of $C(A)$


### problem from page 2 of the text

```{r}
#systems of equations
A <- matrix(c(2,-1,1.5,1,0,-4),nrow=2,byrow = TRUE)

#solution given in the text
x <- c(5,6.5,3)

A%*%x

A
Rank(A)

#RHS
b <- c(8,-7)
```

```{r}
#Is this Homogeneous equations or inhomogeneous equation?
H <- cbind(A,b)
H
rref(H)
Rank(H)
```

```{r}
#unique solution
A <- matrix(c(1,-2,-1,3), nrow =2 , byrow = TRUE)
A
Rank(A)

b <- c(-1,3)

x <- inv(A)%*%b

#get the solution
A%*%x
```

```{r}
#no solution

A <- matrix(c(1,-2,-1,2), nrow =2 , byrow = TRUE)
Rank(A)

b <- c(-1,3)

#how can I check if we can span b or not?

```

```{r}
#many solution

A <- matrix(c(1,-2,-1,2), nrow =2 , byrow = TRUE)
A
Rank(A)

b <- c(-1,1.5)

H <- cbind(A,b)
#how can I check if we can span b or not?
Rank(H)
```

### See Page 5

```{r}
a <- c(1,-2,1,0,0,2,-8,8,-4,5,9,-9)

A <- matrix(a, nrow=3, byrow = TRUE)
Rank(A)
#okay I made mistake, need to get the
#RHS out
b <- A[,4]
A <- A[,c(1,2,3)]
A
Rank(A)
b

x <- inv(A)%*%b

A%*%x
```

## Solution Sets to Linear Systems

### See Page 45

```{r}
a <- c(3,5,-4,-3,-2,4,6,1,-8)

A <- matrix(a, nrow=3, byrow = TRUE)
Rank(A)

b <- c(7,-1,-4)

H <- cbind(A,b)
Rank(H)

rref(H)

B <- A[, c(1,2)]
N <- A[,c(3)]
```

$$\begin{aligned}
\mathbb{A}\vec{x} &= \vec{b}\\
\mathbb{B}\vec{x}_B + \mathbb{N}\vec{x}_N &= \vec{b}\\
\mathbb{B}^{T}\mathbb{B}\vec{x}_B + \mathbb{B}^{T}\mathbb{N}\vec{x}_N &= \mathbb{B}^{T}\vec{b}\\
\mathbb{G}\vec{x}_B + \mathbb{B}^{T}\mathbb{N}\vec{x}_N &= \mathbb{B}^{T}\vec{b}\\
\mathbb{G}^{-1}\mathbb{G}\vec{x}_B + \mathbb{G}^{-1}\mathbb{B}^{T}\mathbb{N}\vec{x}_N &=\mathbb{G}^{-1} \mathbb{B}^{T}\vec{b}\\
\mathbb{I}\vec{x}_B + \mathbb{G}^{-1}\mathbb{B}^{T}\mathbb{N}\vec{x}_N &=\mathbb{G}^{-1} \mathbb{B}^{T}\vec{b}\\
\vec{x}_B + \mathbb{G}^{-1}\mathbb{B}^{T}\mathbb{N}\vec{x}_N &=\mathbb{G}^{-1} \mathbb{B}^{T}\vec{b}
\end{aligned}$$

$$\begin{aligned}
ax + b_1 &= 5\\
d + 5c &= 12
\end{aligned}$$

```{r}
B
N

K <- t(B)%*%B 

#get the solution
x_b <- inv(K)%*%t(B)%*%b  

#check the solution
B%*%x_b

#  
inv(K)%*%t(B)%*%N
4/3
```

Recall that we have selected first two columns vectors as the basis and
their coordiate is given as $-1$ and $2$. For the vector in $\mathbb{N}$, we got 
$\frac{4}{3}$.
```{r}
print(A)
print(x_b)
```

## Additional information shown in rref(A)

- rref(A) tells you the solution to homogeneous systems of equations (See page 43)
```{r}
rref(A)
```

- rref(A) tells you the relationship between the basis and dependent vectors in 
expressing the solution vector [-1,2,0]

$$\begin{aligned}
x_1  \phantom{00} - \phantom{00}\frac{4}{3}x_3 &= -1\\
\phantom{00} x_2 \phantom{000000}  &= \phantom{-}2\\
\phantom{00}\phantom{00}0 &= \phantom{-}0
\end{aligned}$$

## Parametric description of solution sets

- `free variables` act as parameters.  
  - Can anyone define `parameter`?
  
- See the example of parametric description of solution sets

```{r}
r1 <- c(1,6,2,-5,-2,-4)
r2 <- c(0,0,2,-8,-1,3)
r3 <- c(0,0,0,0,1,7)

A <- rbind(r1,r2,r3)

rref(A)
```


$$\begin{aligned}
x_1 &= -6x_2-3x_4\\
x_2 &= \text{free} \\
x_3 &= 5+4x_4 \\
x_4 &= \text{free}\\
x_5 &= 7
\end{aligned}$$

- The above example has $\infty$ number of solutions

## Homogeneous Linear Systems

```{r}
r1 <- c(3,5,-4,0)
r2 <- c(-3,-2,4,0)
r3 <- c(6,1,-8,0)

A <- rbind(r1,r2,r3)

rref(A)
```

- Now rewrite each row as equations and `free variables` needs to go to the RHS.

$\vec{x} = [ x_1, x_2, x_3 ] = [\frac{4}{3}x_3,0,x_3]$

- Now, let's get the constant out.

$\vec{v} = [\frac{4}{3},0,1]$

- I have not yet explained $\vec{p}$ is, but the idea is that solution set (i.e., hyperplane) can be expressed as `parametric vector equation` of the plane that has the following form.

$$\vec{x} = s\vec{p} + t\vec{v}$$

- People say, the solution is in the `parametric vector form`

```{r}
r1 <- c(3,5,-4)
r2 <- c(-3,-2,4)
r3 <- c(6,1,-8)

b <- c(7,-1,-4)

A <- rbind(r1,r2,r3)
Ab <- cbind(A,b)
rref(Ab)
```

$\vec{x} = [ x_1, x_2, x_3 ]^T = [-1,2,0]^T + [\frac{4}{3}x_3,0,x_3]^T = [-1,2,0]^T + x_3[\frac{4}{3},0,1]$

- Recall that $\mathbb{A}\vec{x}=\vec{0}$ has the parametric vector solution.

- The solution to $\mathbb{A}\vec{x}=\vec{b}$ can be found by  shifting the solution to the $\mathbb{A}\vec{x}=\vec{0}$, which is a subspace, by a constant vector $\vec{p}$.  The resulting solution set is `hyperplane` 


## More examples

Given:

$$x_1+6x_2 + 3x_4 = 0$$
$$3x_3 - 4x_4 = 5$$
$$x_5 = 7$$

```{r}
a1 <- c(1,6,0,3,0,0)
a2 <- c(0,0,1,-4,0,5)
a3 <-  c(0,0,0,0,1,7)
a <- c(a1,a2,a3)

A <- matrix(a,nrow=3,byrow=T)
print(A)

print(rref(A))
```

This is parametric description of solution set. Affine subspace or hyperplane.

$$x_1 = -6x_2 -3x_3$$
$$x_2 \text{ is free}$$
$$x_3 = 5 +4x_4$$
$$x_4 \text{ is free}$$
$$x_5 = 7$$

## Network example

```{r}
a1 <- c(1,1,0,0,0,800)
a2 <- c(0,1,-1,1,0,300)
a3 <-  c(0,0,0,1,1,500)
a4 <-  c(1,0,0,0,1,600)
a5 <-  c(0,0,1,0,0,400)
a <- c(a1,a2,a3,a4,a5)

A <- matrix(a,nrow=5,byrow=T)
print(A)
rref(A)

B <- A[,1:4]
b <- A[,c(6)]
print(B)
print(b)
print(Rank(B))
print(Rank(A))
```

```{r}
inv(t(B)%*%B)%*%t(B)%*%b
```

#  More about invertible matrix 

### Given: Suppose $A\in R^{nxn}$ and $A^{-1}$ exist, then the following can be said
- The columns of $A$ is the basis of $R^{n}$
- rank $A$ = n
- $Nul A=$ {$\vec{0}$}
- dim $Nul A$ = 0
- $A^{-1}A=I$
- $AA^{-1}=I$
- The Linear transformation $\vec{x} \mapsto A\vec{x}$ is one-to-one
- $A^T$ is an invertible matrix

- Rouche-Capelli Theorem <
  - Tells you the number of solutions in the systems of equations
  - Alfredo Capelli (5 August 1855 – 28 January 1910)
  
## Change of basis 

Given:  $\vec{y} \notin C(A)$ , and Rank of $A$ = 2, and $\vec{y} \in R^3$

## ```Problem:```  
- Let $\hat{\vec{y}} \subset C(A)$ where $\vec{C}_1 \text{ and } \vec{C}_2$ are the basis of $C(A)$
- Find $\hat{\vec{y}}$ that minimizes $||\vec{y}-\hat{\vec{y}}||$

## Solution:
- let $C$ and $N$ be the matrix that contains the basis of $C(A)$ and $N(A^T)$
- Since: $C\vec{x}=\hat{\vec{y}} \text{   and   } C\vec{x} + N\vec{z} = \vec{y}$
- Simplify the expression
$$C^TC\vec{x} = C^T\vec{y}$$
$$\vec{x} = (C^TC)^{-1}C^T\vec{y}$$

- Then, $C(C^TC)^{-1}C^T\vec{y}=\hat{\vec{y}}$
- $C(C^TC)^{-1}C^T$ is called ```projection matrix```

## Projection matrix


$\vec{y} = P\vec{y} + B\vec{y}$ <br>
$P + B = I$


- where  $P \text{ and } B$ are the ```projection matrices``` for $C(A) \text{ and } N(A^T)$

## Dot product

$$\hat{\vec{y}} = P_{\vec{u}}^{\vec{y}} = \frac{\vec{y}\centerdot\vec{u}}{\vec{u}\centerdot\vec{u}}\vec{u}$$

where <br>
$\vec{y}\centerdot\vec{u} \text{ and } \vec{u}\centerdot\vec{u}$ are scalar quantity.

Projection tells you the ```length``` of the ```projected vector```, $\hat{\vec{y}}$ in terms of the vector that is ```being projected onto``` $\vec{u}$

### Using Projection matrix

```{r}
# y will be projected onto u
y <- matrix(c(7,6),nrow=2)
u <- matrix(c(4,2),nrow=2)
u0 <- matrix(c(16,8),nrow=2)

## using projection matrix
P <- u%*%(solve(t(u)%*%u)%*%t(u))
print(P)

print(P%*%y)
```

### ```Projection formula``` on to $\vec{u}$

```{r}
print(drop((t(y)%*%u)/(t(u)%*%u))*u)
print(drop((t(y)%*%u)/(t(u)%*%u)))
```

Using ```Projection formula``` on to $\vec{u}_0$

```{r}
print(drop((t(y)%*%u0)/(t(u0)%*%u0))*u0)
print(drop((t(y)%*%u0)/(t(u0)%*%u0)))
```

## Orthonormal basis

```{r}
mybasis <- matrix(c(1,2,3,4,5,6),nrow=3)
print(mybasis)
print(orthonormalization(mybasis))
```

```{r}
Z <- (orthonormalization(mybasis)) # z is orthonormal basis of codomain (I called it output space)
```

```{r}
A <- matrix(c(4,3,5,6,8,10,5,12,13),nrow=3, byrow=T)
print(A)

c(Norm(A[1,]),Norm(A[2,]),Norm(A[3,])) #norm of each row vectors in A (i.e., sample)
```

```{r}
B <- A%*%Z
print(B)
print(Z)
```

## Change of basis

```{r}
a <- c(2/3,-2/3,1/3,2/3,1/3,-2/3)
A <- matrix(a,nrow=3,ncol=2,byrow=T)
print(fractions(A))

k <- rnorm(10000,5,5)
myData <- matrix(k,nrow=2,ncol=5000,byrow=T)
c_A <- A%*%myData

C = orthonormalization(A)
GS_A<- cbind(C[,1],C[,2])  #orthonormalized basis spanning C(A)
E <- GS_A%*%myData
```

- $\vec{b} \notin C(A)$ and suppose I want to flip $\vec{b}$ with respect to $C(A)$
- How would you develop the transformation matrix that flips $\vec{b}_0$?

```{r}
b0 <- matrix(c(10,10,10),nrow=3)
B <- cbind(A,b0)
print(rref(B))
```

- Let $D$ transformation matrix that is expressed in terms of the basis spanning $C(A)$ and $N(A^T)$
- Then, 


\begin{equation}
D = \left(
\begin{array}{cccc}
1 & 0 &  0 \\
0 & 1 &  0 \\
0 & 0 &-1 \\
\end{array}
\right)
\end{equation}

- Let $\vec{b}_{1C}$ be the flipped vector whose coordiate is expressed in terms of basis spanning $C(A)$ and $N(A^T)$
- $T\vec{b}_0=\vec{b}_{\text{1standard basis}}$ and $D\vec{b}_{0C} = \vec{b}_{1C}$
- Then, using 
 - $\vec{b}_{0C} = \text{[0 0 1]}^T$ and
 - $\vec{b}_{0} = \text{[10 10 10]}^T$ and
 - $C$ and $D$, we can get $T$
 
### Find $\vec{b}_{\text{1standard basis}}$

```{r}
D <- matrix(c(1,0,0,0,1,0,0,0,-1), nrow=3, byrow=T)
dim(C)
print(D)
T <- C%*%D%*%inv(C)
print(T%*%b0)
```

## Change of basis is useful

- Can you explain why?
- Developing transformation matrix
- Evaluating long term behavior of the transformation matrix
- Understanding how the initial state will evolve over time
- Determining the influence of each basis of $C(A^T)$ to the $C(A)$

## Question

- Explains what happens when you project $\vec{x}$ onto
  - basis that are not orthogonal
    - Try add the projected ones and compare with the original 
  - basis that are orthogonal, but not normal
  - basis that are **orthonormal**
and add the projected vector.
