---
title: "I wonder By Sam"
author: "Bill Chung"
date: '\today'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma)
library(MASS)
library(far)
```


## Introduction

```{r}
print("hello")
```

## Chapter 1

### Section 1.1
```{r}
v1 <- matrix(c(1,-1), nrow=2)
print(v1)
```


```{r}
v1 <- matrix(c(3,4), nrow=2)
print(v1)

r1 <- matrix(c(1,-1,3), nrow=1)
print(r1)
```


```{r}
v1 <- matrix(c(3,4),nrow=2)
print(v1)

v2<-matrix(c(-1.5,2),nrow=2)
print(v2)

print(-2*v2)
```


### Section 1.2

```{r}
v1 <- matrix(c(8,6),nrow=2)
print(v1)

print(t(v1))
```

```{r}
v2 <- matrix(c(4,0),nrow=2)

print(t(v1)%*%v2)
```

### Section 1.3

```{r}
v1 <- matrix(c(8,6),nrow=2)
print(v1)
Norm(v1)
```

```{r}
v2 <- matrix(c(4,0),nrow=2)
print(v2)
```

```{r}
print(t(v1)%*%v2/(t(v1)%*%v1))

print(t(v1)%*%v2/Norm(v1)^2)

```


\break

# Chapter 2

## Section 2.1

```{r}
v1 <- matrix(c(1,-1), nrow =2)
v2 <- matrix(c(2,1), nrow =2)
A <- cbind(v1,v2)
print(A)
```

```{r}
dim(A)
```


## Section 2.2

```{r}
v1 <- matrix(c(1,1), nrow =2)
v2 <- matrix(c(1,-1), nrow =2)
v3 <- matrix(c(1,3), nrow =2)

A <- cbind(v1,v2,v3)
print(A)

x <- matrix(c(2,3,1.5),nrow=3)
print(x)

A%*%x
```

## Section 2.3

```{r}
v1 <- matrix(c(1,-1), nrow =2)
v2 <- matrix(c(2,-1), nrow =2)
v3 <- matrix(c(-2,-1), nrow =2)
A <- cbind(v1,v2,v3)
print(A)

Rank(A)
```

```{r}
v1 <- matrix(c(1,-1), nrow =2)
v2 <- matrix(c(2,-1), nrow =2)
v3 <- matrix(c(-2,-3), nrow =2)
A <- cbind(v1,v2,v3)
print(A)

print(rref(A))
```

### problem 1

```{r}
r1 <- matrix(c(-3,1,-1,1,4), nrow =1)
r2 <- matrix(c(1,-2,2,3,1), nrow =1)
r3 <- matrix(c(-3,1,-1,1,4), nrow =1)
A <- rbind(r1,r2,r3)
print(A)
rref(A)
```

### problem 2

```{r}
r1 <- matrix(c(-3,-6,-1,-2,4), nrow =1)
r2 <- matrix(c(1,2,2,4,1), nrow =1)
r3 <- matrix(c(2,4,5,10,7), nrow =1)
A <- rbind(r1,r2,r3)
print(A)
rref(A)
```

### problem 3

```{r}
r1 <- matrix(c(-3,-1,4), nrow =1)
r2 <- matrix(c(1,2,1), nrow =1)
r3 <- matrix(c(2,5,7), nrow =1)
A <- rbind(r1,r2,r3)
print(A)
rref(A)
```

### problem 4

```{r}
r1 <- matrix(c(-3,-1), nrow =1)
r2 <- matrix(c(1,2), nrow =1)
r3 <- matrix(c(2,5), nrow =1)
A <- rbind(r1,r2,r3)
print(A)
print(rref(A))
```

### problem 5

```{r}
r1 <- matrix(c(-3), nrow =1)
r2 <- matrix(c(1), nrow =1)
r3 <- matrix(c(2), nrow =1)
A <- rbind(r1,r2,r3)
print(A)
rref(A)
```


## Section 2.4

```{r}
v1 <- matrix(c(1,-1), nrow =2)
v2 <- matrix(c(2,-1), nrow =2)
v3 <- matrix(c(-2,-3), nrow =2)

A <- cbind(v1,v2,v3)
print(A)

rref(A)
```

```{r}
B <- A[,c(1,2)]
print(B)
```

```{r}
D <- A[,-c(1,2)]
print(D)
```

```{r}
x <- matrix(c(2,3,7), nrow=3)
print(x)
```

```{r}
xB <- matrix(c(2,3), nrow=2)
print(xB)
```

```{r}
xD <- matrix(c(7), nrow=1)
print(xD)
```

```{r}
b <- A%*%x
print(b)
```

```{r}
print(B%*%xB + D%*%xD)
```

### problem 1

```{r}
v1 <- matrix(c(1,-1), nrow =2)
v2 <- matrix(c(2,-1), nrow =2)
v3 <- matrix(c(-2,-3), nrow =2)

A <- cbind(v1,v2,v3)
print(A)

x<- matrix(c(2,3,7), nrow =3)
print(x)

b <- A%*%x
print(b)
print(rref(A))

B<- A[,c(1,2)]
print(B)

D<- A[,-c(1,2)]
print(D)

```

### problem 2


```{r}
r1 <- matrix(c(-3,1,-1,1,4), nrow =1)
r2 <- matrix(c(1,-2,2,3,1), nrow =1)
r3 <- matrix(c(2,-4,5,8,7), nrow =1)
A <- rbind(r1,r2,r3)
print(A)

x<- matrix(c(2,3,7,-1,4),nrow=5)

b <- A%*%x
print(b)
rref(A)
```

```{r}
B <- A[,c(1,2,3)]
D <- A[,-c(1,2,3)]
xB <- matrix(c(2,3,7),nrow=3)
xD <- matrix(c(-1,4),nrow=2)
print(B%*%xB + D%*%xD)
```

### problem 3


```{r}
r1 <- matrix(c(-3,-6,-1,-2,4), nrow =1)
r2 <- matrix(c(1,2,2,4,1), nrow =1)
r3 <- matrix(c(2,4,5,10,7), nrow =1)
A <- rbind(r1,r2,r3)
print(A)

x<- matrix(c(2,4,7,3,1),nrow=5)

b <- A%*%x
print(b)
rref(A)
```

```{r}
B <- A[,c(1,3,5)]
D <- A[,-c(1,3,5)]
xB <- matrix(c(2,7,1),nrow=3)
xD <- matrix(c(4,3),nrow=2)
print(B%*%xB + D%*%xD)
```

## Section 2.5

```{r}
r1 <- matrix(c(-1,1,1,3,-1), nrow =1)
r2 <- matrix(c(3,-2,-1,-7,5), nrow =1)
r3 <- matrix(c(2,1,4,0,8), nrow =1)
A <- rbind(r1,r2,r3)
print(A)
```


```{r}
r1 <- matrix(c(1,3,2,-4,3), nrow =1)
r2 <- matrix(c(1,-2,3,-1,2), nrow =1)
F <- rbind(r1,r2)
print(F)
```

```{r}
A%*%t(F)
```

### Problem 1

```{r}
r1 <- matrix(c(-1,1,1,3,-1), nrow =1)
r2 <- matrix(c(3,-2,-1,-7,5), nrow =1)
r3 <- matrix(c(2,1,4,0,8), nrow =1)
A <- rbind(r1,r2,r3)
print(A)
Rank(A)

GA <- t(A)%*%A
print(GA)
inv(GA)
```


### Problem 2

```{r}
r1 <- matrix(c(-1,1), nrow =1)
r2 <- matrix(c(3,-2), nrow =1)
r3 <- matrix(c(2,1), nrow =1)
A <- rbind(r1,r2,r3)
print(A)
Rank(A)

GA <- t(A)%*%A
print(GA)
print(round(inv(GA),2))
```

### Problem 3

```{r}
r1 <- matrix(c(-1,1,1), nrow =1)
r2 <- matrix(c(3,-2,-1), nrow =1)
r3 <- matrix(c(2,1,4), nrow =1)
A <- rbind(r1,r2,r3)
print(A)
Rank(A)

GA <- t(A)%*%A
print(GA)

inv(GA)
```


## Section 2.6

```{r}
diag(c(-4,0,1))

diag(c(5,3,1))

diag(c(6,2,1))
```


```{r}
r1 <- matrix(c(2,4,2), nrow =1)
r2 <- matrix(c(0,3,1), nrow =1)
r3 <- matrix(c(2,5,2), nrow =1)
P <- rbind(r1,r2,r3)
print(P)
print(inv(P))
print(inv(P)%*%P)
```


```{r}
r1 <- matrix(c(2,0), nrow =1)
r2 <- matrix(c(0,1), nrow =1)
r3 <- matrix(c(2,1), nrow =1)
B <- rbind(r1,r2,r3)
print(B)
GB <- t(B)%*%B
print(GB)
print(round(inv(GB),2))
print(inv(GB)%*%GB)
```

### Problem 1

```{r}
r1 <- matrix(c(2,0,4,4,2), nrow =1)
r2 <- matrix(c(0,1,2,3,1), nrow =1)
r3 <- matrix(c(2,1,3,5,2), nrow =1)
A <- rbind(r1,r2,r3)
print(A)
print(round(rref(A),2))

B <- A[,c(1,2,3)]
print(B)
print(round(inv(B)%*%B),2)
```


### Problem 2

```{r}
v1 <- matrix(c(2,0,2,4,5), nrow =5)
v2 <- matrix(c(0,1,1,3,4), nrow =5)
v3 <- matrix(c(4,2,4,7,7), nrow =5)
A <- cbind(v1,v2,v3)
print(A)
Rank(A)

GA <- t(A)%*%A
print(GA)

print(round(inv(GA)%*%GA),2)
```

### Problem 3

```{r}
v1 <- matrix(c(2,0,2), nrow =3)
v2 <- matrix(c(0,1,1), nrow =3)
v3 <- matrix(c(4,1,5), nrow =3)
A <- cbind(v1,v2,v3)
print(A)
print(rref(A))

B <- A[,c(1,2)]
print(B)

GB <- t(B)%*%B
print(GB)

print(inv(GB)%*%GB)
```

### Problem 4

```{r}
v1 <- matrix(c(2,0,2), nrow =3)
v2 <- matrix(c(0,1,1), nrow =3)
v3 <- matrix(c(4,2,4), nrow =3)
A <- cbind(v1,v2,v3)
print(A)
print(rref(A))
I <- inv(A)%*%A
print(I)
```

### Problem 5

```{r}
v1 <- matrix(c(2,0,2), nrow =3)
A <- v1

GA <- t(A)%*%A
print(GA)
print(round(inv(GA)%*%GA),2)
```


\break

# Chapter 3

## Section 3.1

```{r}
A <- matrix(c(1,1,1,1,-1,-3), nrow = 2, byrow=T)
print(A)
dim(A)
dim(A)[1]
dim(A)[2]
```

## Section 3.3


```{r}
r1 <- matrix(c(1.1,1.1,3.3,1.1), nrow =1)
r2 <- matrix(c(-1.2,3.3,5.4,7.8), nrow =1)
r3 <- matrix(c(1.3,2.2,5.7,3.1), nrow =1)
A <- rbind(r1,r2,r3)
print(A)

print(rref(A))
Rank(A)

B <- A[,c(1,2)]
print(B)

D <- A[,-c(1,2)]
print(D)

GB <- t(B)%*%B
print(GB)

print(-inv(GB)%*%t(B)%*%D)

I <- diag(2)
print(I)

print(rbind(-inv(GB)%*%(t(B)%*%D),I))

N <- rbind(-inv(GB)%*%(t(B)%*%D),I)
print(round(A%*%N,2))

```



```{r}
print(A)
AT <- t(A)
print(AT)
print(rref(t(A)))
RA <- AT[,c(1,2)]
print(RA)
```


## Section 3.5

```{r}
library(far)

v1 <- matrix(c(1,2,3), nrow =3)
v2 <- matrix(c(4,5,6), nrow =3)
v3 <- matrix(c(13,2,-4), nrow =3)
A <- cbind(v1,v2,v3)
print(A)
Rank(A)

OA <- orthonormalization(A)
print(round(OA,2))

Rank(OA)

Norm(OA[,1])

#between column vectors
print(round(t(OA[,1])%*%OA[,2],2))
print(round(t(OA[,1])%*%OA[,3],2))

#between row vectors
print(round(t(OA[1,])%*%OA[2,],2))
print(round(t(OA[1,])%*%OA[3,],2))
```


```{r}
v1 <- matrix(c(1,2,-3), nrow =3)
v2 <- matrix(c(-4,2,1), nrow =3)
v3 <- matrix(c(11,2,-11), nrow =3)
A <- cbind(v1,v2,v3)
print(A)

print(rref(A))

B <- A[,c(1,2)]
print(B)

OB <- orthonormalization(B)
print(round(OB,2))
```


### Problem 1


```{r}
v1 <- matrix(c(1,2,-3), nrow =3)
v2 <- matrix(c(-4,2,1), nrow =3)
v3 <- matrix(c(-3,4,-2), nrow =3)
A <- cbind(v1,v2,v3)
print(A)
print(rref(A))

B <- A[,c(1,2)]

OB <- orthonormalization(B)
print(round(OB,2))

#basis spanning column space of A
CA <- OB[,c(1,2)]
print(round(CA,2))

#basis spanning left nullspace of A
NAT <- OB[,c(3)]
print(round(NAT,2))

```

### Problem 2

```{r}
r1 <- matrix(c(1,-4,-3,-3,23), nrow =1)
r2 <- matrix(c(2,2,4,4,-1), nrow =1)
r3 <- matrix(c(-3,1,-2,-2,-14), nrow =1)
A <- rbind(r1,r2,r3)
print(A)
print(rref(A))

B <- A[,c(1,2)]

OB <- orthonormalization(B)
print(round(OB,2))

#basis spanning column space of A
CA <- OB[,c(1,2)]
print(round(CA,2))

#basis spanning left nullspace of A
NAT <- OB[,-c(1,2)]
print(round(NAT,2))
```

\break 

# Chapter 4

## Section 4.1

```{r}
v1 <- matrix(c(1,-1,1), nrow =3)
v2 <- matrix(c(1,3,2), nrow =3)
v3 <- matrix(c(2,3,4), nrow =3)
A <- cbind(v1,v2,v3)
print(A)

Rank(A)

b <- matrix(c(1,4,-4))
print(b)

Rank(b)

x <-inv(A)%*%b
print(round(x,2))

print(A%*%x)
```


## Section 4.2

```{r}
v1 <- matrix(c(1,-1,1), nrow =3)
v2 <- matrix(c(1,3,2), nrow =3)
v3 <- matrix(c(3,5,5), nrow =3)
A <- cbind(v1,v2,v3)
print(A)

b <- matrix(c(1,4,-4))
print(b)

Ab <- cbind(A,b)

Rank(A)
Rank(Ab)

print(rref(A))

B <- A[,c(1,2)]
GB <- t(B)%*%B
xB <- inv(GB)%*%t(B)%*%b
print(round(xB,2))

b_hat <- B%*%xB
print(round(b_hat,2))

x <- rbind(xB,0)
print(round(x,2))

print(round(A%*%x,2))
```
## Section 4.3

### Problem 1

```{r}
r1 <- matrix(c(1,1,2,2,3), nrow =1)
r2 <- matrix(c(-1,3,3,-1,5), nrow =1)
r3 <- matrix(c(1,2,4,3,5), nrow =1)
A <- rbind(r1,r2,r3)
print(A)

b <- matrix(c(1,4,-4))
print(b)

######################################
#STEP 1
######################################

print(rref(A))

Ab <- cbind(A,b)
print(round(rref(Ab),2))

B<- A[,c(1,2,3)]
D<- A[,-c(1,2,3)]
print(B)
print(D)

GB <- t(B)%*%B
print(GB)

xB <- inv(GB)%*%t(B)%*%b
print(round(xB,2))

x_zero <- rbind(xB[1],xB[2],xB[3],0,0)
print(round(x_zero,2))

###########################################
# STEP 2
###########################################
N <- rbind(-inv(GB)%*%t(B)%*%D, diag(2))
print(round(N,0))

# point 1
x_D <- matrix(c(-3,2),nrow=2)
print(x_D)

x_1 <- x_zero - N%*%x_D
print(round(x_1,2))

print(A%*%x_1)

# point 2
x_D <- matrix(c(1,5),nrow=2)
print(x_D)

x_2 <- x_zero - N%*%x_D
print(round(x_2,2))

print(A%*%x_2)
```


\break 

# Chapter 5

## Section 5.1


```{r}
r1 <- matrix(c(-2,4,2), nrow =1)
r2 <- matrix(c(-4.8,8.4,1.6), nrow =1)
r3 <- matrix(c(-4.8,2.4,5.6), nrow =1)
A <- rbind(r1,r2,r3)
print(A)

eigen(A)

eigen(A)$values
round(eigen(A)$vectors,2)
```

### Problem 1

```{r}
r1 <- matrix(c(1,4,7), nrow =1)
r2 <- matrix(c(2,5,8), nrow =1)
r3 <- matrix(c(3,6,9), nrow =1)
A <- rbind(r1,r2,r3)
print(A)

Rank(A)

L <- diag(eigen(A)$values)
print(round(L,2))

E <- eigen(A)$vectors
print(round(E,2))

Rank(E)

```


### Problem 2

```{r}
r1 <- matrix(c(1,4,2), nrow =1)
r2 <- matrix(c(5,5,8), nrow =1)
r3 <- matrix(c(-3,69,9), nrow =1)
A <- rbind(r1,r2,r3)
print(A)

Rank(A)

L <- diag(eigen(A)$values)
print(round(L,2))

E <- eigen(A)$vectors
print(round(E,2))

Rank(E)
```


### Problem 3

```{r}
r1 <- matrix(c(1,2,2), nrow =1)
r2 <- matrix(c(2,4,8), nrow =1)
r3 <- matrix(c(3,6,9), nrow =1)
A <- rbind(r1,r2,r3)
print(A)

Rank(A)

L <- diag(eigen(A)$values)
print(round(L,2))

E <- eigen(A)$vectors
print(round(E,2))

Rank(E)
```

### Problem 4

```{r}
r1 <- matrix(c(0.8,-0.6,0), nrow =1)
r2 <- matrix(c(0.6,0.8,0), nrow =1)
r3 <- matrix(c(3,3,1), nrow =1)
A <- rbind(r1,r2,r3)
print(A)

Rank(A)

L <- diag(eigen(A)$values)
print(round(L,2))

E <- eigen(A)$vectors
print(round(E,2))


```



## Section 5.2

```{r}
r1 <- matrix(c(-2,4,2), nrow =1)
r2 <- matrix(c(-4.8,8.4,1.6), nrow =1)
r3 <- matrix(c(-4.8,2.4,5.6), nrow =1)
A <- rbind(r1,r2,r3)
print(A)

Rank(A)

Q<- A - diag(eigen(A)$values[1],3)

print(Q)

Rank(Q)
```


## Section 5.3

```{r}
r1 <- matrix(c(1,4,2), nrow =1)
r2 <- matrix(c(5,5,8), nrow =1)
r3 <- matrix(c(-3,69,9), nrow =1)
A <- rbind(r1,r2,r3)
print(A)

L <- diag(eigen(A)$values)
print(round(L,2))

E <- eigen(A)$vectors
print(round(E,2))

```


```{r}
A <- matrix(c(0,0,2.4,1.2), nrow=2)
print(A)

#check eigendecomposition
eigen(A)

#create L
L <- diag(eigen(A)$values)
L

#check if L is singular
inv(L)
```


## Section 5.4

### Problem 1

```{r}
r1 <- matrix(c(1,2,2), nrow =1)
r2 <- matrix(c(2,4,8), nrow =1)
r3 <- matrix(c(3,6,9), nrow =1)
A <- rbind(r1,r2,r3)
print(A)

E <- eigen(A)$vectors
print(round(E,2))

x<- matrix(c(2,4,8),nrow=3)
print(x)

h<- inv(E)%*%x

h[1]%*%E[,1] + h[2]%*%E[,2] + h[3]%*%E[,3]
```


### Problem 2

```{r}
L <- diag(c(1.2,0.7))
e1 <- c(2,1)
e2 <- c(1,0)
E <- cbind(e1,e2)
A <- E%*%L%*%inv(E)
print(A)

```

## Section 5.5

### Problem 1

```{r}
r1 <- matrix(c(1,2,2), nrow =1)
r2 <- matrix(c(2,4,8), nrow =1)
r3 <- matrix(c(3,6,9), nrow =1)
A <- rbind(r1,r2,r3)
print(A)

Rank(A)

GA <- t(A)%*%A
print(GA)

L <- round(diag(eigen(GA)$values,3),1)
print(L)

E <- round(eigen(GA)$vectors,3)
E

t(E)

round(inv(E),3)
```

### Problem 2


```{r}
r1 <- matrix(c(1,2,-1), nrow =1)
r2 <- matrix(c(2,2,3), nrow =1)
r3 <- matrix(c(0,3,2), nrow =1)
A <- rbind(r1,r2,r3)
print(A)

Rank(A)

GA <- t(A)%*%A
print(GA)

L <- round(diag(eigen(GA)$values,2),1)
print(L)

E <- round(eigen(GA)$vectors,2)
E

round(t(E),2)

round(inv(E),2)
```

## Section 5.6

### Problem 1

```{r}
M <- matrix(c(0.7,0.25,0.3,0.75), 
            nrow=2, byrow = T)

print(M)

E <- eigen(M)$vectors
print(round(eigen(M)$vectors,2))

print(eigen(M)$values)

x0 <- matrix(c(0.6,0.4),nrow=2)
print(x0)  

p1 <- E[,1]/sum(E[,1])
print(round(p1,2))

```


## Section 5.7


```{r}
Q <- matrix(c(0,0.5,0,0.5,0,0.5,0,0.5,0), 
            nrow = 3, byrow =T)
print(Q)

N <- inv(diag(3)-Q)
print(N)

R <- matrix(c(0.5,0,0,0,0,0.5), nrow=3, byrow = T)
print(R)

```

### Problem 1

```{r}
N <- inv(diag(3)-Q)
print(N)
```

### Problem 2

```{r}
c1 <- matrix(c(1,1,1), nrow=3)
print(N%*%c1)
```

### Problem 3

```{r}
print(N%*%R)
```

\break

# Chapter 6

## Section 6.1

```{r}
A <- matrix(c(1,2,-2,-1,-1,-3), 
            nrow=2, byrow = TRUE)

print(A)
Rank(A)

S <- diag(svd(A)$d)
print(round(S,2))


U <- svd(A)$u
print(round(U,2))

print(round((svd(A)$u)%*%S),2)

V <- svd(A)$v
print(round(V,2))

print(round(A%*%(svd(A)$v)),2)

U%*%S%*%t(V)
```


### Problem 1

```{r}
A <- matrix(c(1,1,2,1,-1,3), 
            nrow=2, byrow = TRUE)

print(A)

S <- diag(svd(A)$d)
print(round(S,2))

U <- svd(A)$u
print(round(U,2))

V <- svd(A)$v
print(round(V,2))

print(round(A%*%(svd(A)$v),0))

print(round(svd(A)$u%*%S,0))

print(U%*%S%*%t(V))
```

### Problem 2

```{r}
A <- matrix(c(1,4,7,1,1,9), 
            nrow=3, byrow = TRUE)

print(A)

S <- diag(svd(A)$d)
print(round(S,2))


U <- svd(A)$u
print(round(U,2))

V <- svd(A)$v
print(round(V,2))

print(U%*%S%*%t(V))

print(U%*%S%*%inv(V))
```

## Section 6.2

### Problem 1

```{r}
A <- matrix(c(1,2,-2,-1,-1,-3), 
            nrow=2, byrow = TRUE)

print(A)

S <- diag(svd(A)$d)
print(round(S,2))


U <- svd(A)$u
print(round(U,2))

V <- svd(A)$v
print(round(V,2))


A_inv <- V%*%inv(S)%*%inv(U)
print(round(A_inv,2))

result <- A%*%A_inv
print(round(result,2))

```


### Problem 2

```{r}
A <- matrix(c(1,1,2,1,-1,3), 
            nrow=2, byrow = TRUE)

print(A)

S <- diag(svd(A)$d)
print(round(S,2))


U <- svd(A)$u
print(round(U,2))

V <- svd(A)$v
print(round(V,2))


A_inv <- V%*%inv(S)%*%inv(U)
print(round(A_inv,2))

result <- A%*%A_inv
print(round(result,2))

```

## Section 6.3

```{r}
v1 <- matrix(c(5,2),nrow=2)
v2 <- matrix(c(10,0),nrow=2)
v3 <- matrix(c(0,1),nrow=2)

#project onto v2
print(t(v1)%*%v2/(t(v2)%*%v2))


#project onto v2
print(t(v1)%*%v3/(t(v3)%*%v3))
```

- How to normalize a vector

```{r}
#how to normalize v2
nv2 <- v2 / sqrt(sum(v2^2)) 
print(nv2)
print(Norm(nv2))

#how to normalize v2
nv3 <- v3 / sqrt(sum(v3^2)) 
print(nv3)
print(Norm(nv3))

```

- Projecting onto normalized vector

```{r}

#project onto v2
print(t(v1)%*%nv2/(t(nv2)%*%nv2))


#project onto v2
print(t(v1)%*%nv3/(t(nv3)%*%nv3))
```

- orthogonal matrix and a vector multiplication

```{r}
A <- matrix(c(1,2,3,-1,3,2), nrow = 3)
A
Rank(A)

OA <- orthonormalization(A)
OA

x <- matrix(c(2,1,5),nrow = 3)
x

b <- OA%*%x
b

t(OA)%*%b

t(b)%*%OA

```

- principal component analysis 

```{r}
A <- matrix(c(1,2,3,4,5,
              0.2,0.8,0.7,1.1,1.7), 
              nrow=2, byrow = TRUE)

print(A)

S <- diag(svd(A)$d)
print(round(S,2))


U <- svd(A)$u
print(round(U,2))

V <- svd(A)$v
print(round(V,2))

P <- t(A)%*%U
print(round(P,2))



```

